Disentangling patterns of gene expression in high grade serous ovarian cancer by matrix factorization

### 2020-08-13 MSc_Dissertation_B156476.tex
Organised appendices

\usepackage{caption}
\captionsetup[figure]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\captionsetup[table]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{placeins}

@@

GEA against the GO results in, for each of the 11 metagenes, a table of GO terms with a list of the candidate genes which have an annotated association to that term; this table can be found in the appendix, table \ref{tab-go-enrichment-results}.
To take full advantage of the hierarchical nature of the GO, results are presented as lineage maps in figures \ref{fig-NMF_1-go-lineage} to \ref{fig-PCA-3-3-go-lineage}.  
The contributing candidate genes for all significantly enriched terms are shown beneath each figure.  
High-level terms at depth less than 3 were removed, since these are generic (e.g. ``regulation of biological process") and so uninteresting.   
One component -- ICA-4 -- was problematic in that 49 GO terms (with depth $\geq 3$) were identified as significant.  
For this component it was necessary to limit the graphic to the 12 terms having the largest number of associated candidate genes.

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5,draft]{\plotspath{BoxAndScatters/genomic_feature_scatters_TCGA_AOCS.pdf}} 
\caption{Grid of scatter plots to visualise the correlation between metasamples and genomic features.  In the case of WGD which has binary values (0, 1), jitter is applied for visual effect only.   Above each plot is shown a correlation coefficient $r$ and associated p-value, highlighted with '***' where $p < 0.01$.  In the case of WGD , the Point-Biserial correlation is used; for all others Pearson's $r$ is used.}
\label{fig-genomic-scatters}
\end{center}
\end{figure}

@@

\small
\bibliography{OvarianCancer}
\bibliographystyle{ieeetr}

\section{Appendices}
\subsection{Additional figures and plots}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from NMF facorizations on the AOCS dataset, comparing fixed and bootstrap sampling}
\label{fig-AOCS-nmf-fixed-vs-bootstrap}
\end{center}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_TCGA.pdf}} 
\caption{Kaplan-Meier plots for each metasample component, stratified at the median value, for TCGA $\rightarrow$ TCGA for overall survival (OS) case.  Hazard ratio and p-value is shown for each case.  The final plot (bottom right) is unstratified overall survival}
\label{fig-kaplan-meier-os-TCGA-TCGA}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_pfs_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for progression free survival (PFS).  }
\label{fig-kaplan-meier-pfs-TCGA-AOCS}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.7]{\plotspath{HeatMaps/clustered_heatmap_TCGA_AOCS.pdf}} 
\caption{Heatmap of the metasamples matrix for the AOCS dataset, computed from metagenes factorized from the TCGA dataset.   Several patient metadata variables are shown in columns to the left.}
\label{fig-heatmap-TCGA-AOCS}
\end{center}
\end{figure}

\subsection{Additional plots relating to the batch effect investigation}
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{BOTH_AOCS_TCGA/FactorClustering/multiple_single_factors_scatter_BOTH_NMF_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{BOTH_AOCS_TCGA/FactorClustering/multiple_single_factors_scatter_BOTH_ICA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{BOTH_AOCS_TCGA/FactorClustering/multiple_single_factors_scatter_BOTH_PCA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=0.55]{\plotspath{/BOTH_AOCS_TCGA/FactorClustering/silhouette_plots_BOTH_2_3_4_5_6_7_8_9_10_bootstrap.pdf}} 
\caption{Metagene clustering for all three methods applied to the combined N=80+374 AOCS + TCGA dataset with bootstrap sampling, over a range of factorization ranks.  The silhouette scores are also plotted (bottom).  }
\label{fig-BOTH-all-bootstrap}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{BoxAndScatters/genomic_feature_scatters_BOTH_AOCS.pdf}} 
\caption{Grid of scatter plots to visualise the correlation between metasamples derived from the metagenes of the combined AOCS + TCGA datasets, and genomic features of the AOCS.}
\label{fig-genomic-scatters-BOTH-AOCS}
\end{center}
\end{figure}

\subsection{Gene enrichment raw results}
\label{tab-go-enrichment-results}
TODO

\label{sec-software-versions}

>>>>

\usepackage{caption}
\captionsetup[figure]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\captionsetup[table]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{placeins}
\usepackage{longtable}

@@

GEA against the GO results in, for each of the 11 metagenes, a table of GO terms with a list of the candidate genes which have an annotated association to that term. See appendix \ref{tab-go-enrichment-results} to access the full table.
To take full advantage of the hierarchical nature of the GO, results are presented as lineage maps in figures \ref{fig-NMF_1-go-lineage} to \ref{fig-PCA-3-3-go-lineage}.  
The contributing candidate genes for all significantly enriched terms are shown beneath each figure.  
High-level terms at depth less than 3 were removed, since these are generic (e.g. ``regulation of biological process") and so uninteresting.   
One component -- ICA-4 -- was problematic in that 49 GO terms (with depth $\geq 3$) were identified as significant.  
For this component it was necessary to limit the graphic to the 12 terms having the largest number of associated candidate genes.

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.4]{\plotspath{BoxAndScatters/genomic_feature_scatters_TCGA_AOCS.pdf}} 
\caption{Grid of scatter plots to visualise the correlation between metasamples and genomic features.  In the case of WGD which has binary values (0, 1), jitter is applied for visual effect only.   Above each plot is shown a correlation coefficient $r$ and associated p-value, highlighted with '***' where $p < 0.01$.  In the case of WGD , the Point-Biserial correlation is used; for all others Pearson's $r$ is used.}
\label{fig-genomic-scatters}
\end{center}
\end{figure}

@@

\small
\bibliography{OvarianCancer}
\bibliographystyle{ieeetr}
\normalsize

\newpage
\section{Appendices}
\subsection{Metagene clustering -- additional figures}
\def\S{0.35}
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from NMF facorizations on the AOCS dataset, comparing fixed and bootstrap sampling}
\label{fig-AOCS-nmf-fixed-vs-bootstrap}
\end{center}
\end{figure}

@@

\FloatBarrier
\subsection{Survival analysis -- additional figures}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.4]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_TCGA.pdf}} 
\caption{Kaplan-Meier plots for each metasample component, stratified at the median value, for TCGA $\rightarrow$ TCGA for overall survival (OS) case.  Hazard ratio and p-value is shown for each case.  The final plot (bottom right) is unstratified overall survival}
\label{fig-kaplan-meier-os-TCGA-TCGA}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.4]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_pfs_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for progression free survival (PFS).  }
\label{fig-kaplan-meier-pfs-TCGA-AOCS}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{HeatMaps/clustered_heatmap_TCGA_AOCS.pdf}} 
\caption{Heatmap of the metasamples matrix for the AOCS dataset, computed from metagenes factorized from the TCGA dataset.   Several patient metadata variables are shown in columns to the left.}
\label{fig-heatmap-TCGA-AOCS}
\end{center}
\end{figure}

\subsection{Batch effect investigation -- Additional figures}
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{BOTH_AOCS_TCGA/FactorClustering/multiple_single_factors_scatter_BOTH_NMF_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{BOTH_AOCS_TCGA/FactorClustering/multiple_single_factors_scatter_BOTH_ICA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{BOTH_AOCS_TCGA/FactorClustering/multiple_single_factors_scatter_BOTH_PCA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=0.55]{\plotspath{/BOTH_AOCS_TCGA/FactorClustering/silhouette_plots_BOTH_2_3_4_5_6_7_8_9_10_bootstrap.pdf}} 
\caption{Metagene clustering for all three methods applied to the combined N=80+374 AOCS + TCGA dataset with bootstrap sampling, over a range of factorization ranks.  The silhouette scores are also plotted (bottom).  }
\label{fig-BOTH-all-bootstrap}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.4]{\plotspath{BoxAndScatters/genomic_feature_scatters_BOTH_AOCS.pdf}} 
\caption{Grid of scatter plots to visualise the correlation between metasamples derived from the metagenes of the combined AOCS + TCGA datasets, and genomic features of the AOCS.}
\label{fig-genomic-scatters-BOTH-AOCS}
\end{center}
\end{figure}

\subsection{Access to gene enrichment analysis table}
\label{tab-go-enrichment-results}

The table of GEA results is too large to conveniently include here.   If required it can be obtained from my github repository at \url{https://github.com/ipoole/HgsocTromics/blob/master/Results/combined_gea_fdr.tsv}.
